### 2016-11-22 doc/paper/paper.tex
bugfix in the example original code snippet (thanks, Luis!), line-number references.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[]{algorithm2e}
\usepackage{changepage}
\usepackage{hyperref}
\definecolor{mygreen}{rgb}{0.1,0.5,0.1}
\definecolor{myblue}{rgb}{0.1,0.1,0.7}

\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\lstset{language=C,
  escapechar=^,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\color{myblue}\ttfamily,
  stringstyle=\color{mygreen}\ttfamily,
  commentstyle=\color{gray}\ttfamily,
  keepspaces=true,
  columns=flexible,
  tabsize=2,
  otherkeywords={bool},
  numbers=left,
  numbersep=1pt,
  numberstyle=\tiny\color{gray}
}

\newcommand{\code}[1]{\lstinline{#1}}
\newcommand{\vignat}{{\sc VigNAT}\xspace}
\newcommand{\vignata}{{\sc VigNAT$\alpha$}\xspace}
\newcommand{\vignatb}{{\sc VigNAT$\beta$}\xspace}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

\newcounter{solOverviewSteps}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Vigor: Toward Carrier-Grade Software Network Functions}

%\author{
%    {\rm Arseniy Zaostrovnykh}\\
%    EPFL
%    \and
%    {\rm Katerina Argyraki}\\
%    EPFL
%    \and
%    {\rm George Candea}\\
%    EPFL
%}
\author{
    {\rm Paper \#113}\\
}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
The objective of this work is to enable developers to productively write
software-based data plane network functions that simultaneously meet the highest
standards of reliability, security, and performance. We present a verification
approach for formally proving properties of stateful network functions, in a way
that is practical both in terms of developer effort and the performance of the
resulting data plane software. Unlike alternative approaches that require
drastic changes in the way network functions are developed, our approach allows
the use of C/C++ together with popular frameworks like DPDK. To the best of our
knowledge, our work is the first to both be practical and formally verify
stateful network functions. We introduce a new ``speculative'' form of symbolic
execution, which achieves scalability to real-world network functions by
leveraging insights related to the structure and cyclomatic complexity of data
plane software. This technique dovetails with a new library of verified
components that are commonly used in network functions. We illustrate our
approach by developing a NAT box that is verified to never crash, to not leak
memory, be free of several classes of security vulnerabilities, and whose
performance is on par with other NATs.
% \footnote{Disclaimer: we use acronym NAT in this paper to refer to a more
% narrow kind - Dynamic Source NAPT (network address-port
% translation~\cite{rfc3022}).}.
We view this as an important milestone on the road to complete verification of all software-based network functions.

\section{Introduction}

Product development in the telecommunication industry has traditionally followed
rigorous standards for quality, security, and protocol adherence, with the term
carrier-grade designating network equipment that meets these high standards.
While this model worked well in the past, it typically entailed reliance on
proprietary hardware and long product cycles, which slowed progress. To meet the
growing demand for new network functionality and faster deployment cycles,
network service providers started turning to software.

In recent years, new platforms emerged for developing network functionality in
software~\cite{kohler2000click,bosshart2014p4,mccanne1993bsd,song2015unified,shah2003np,shahbaz2016pisces},
enabling the rapid prototyping and deployment of new routing
protocols~\cite{sivaraman2016packet}, QoS
policies~\cite{sivaraman2016programmable}, load balancing
algorithms~\cite{zhang2013steering}, intrusion prevention
systems~\cite{xing2014sdnips}, etc. The agility that software engenders
determined even traditional hardware-based network device manufacturers to
invest in software-based approaches~\cite{ciscodevelopssdn}.

But this raises the question of how to develop carrier-grade network functions
(NF) in software? Since NFs are often part of critical infrastructure, they must
simultaneously be secure, reliable, and high-performance, i.e., be
carrier-grade. The Achilles' heel of software has always been the (perceived)
lack of reliability and security, particularly when it is written in C. This has
led researchers to advocate using safe languages~\cite{clipsham2015safe,
  kourtis2015intelligent, fowler2014verified} or relatively heavyweight
approaches~\cite{madhavapeddy2007melange, madhavapeddy2005splat,
  musuvathi2004model, alur2001verifying, ennals2004linear, ridge2008rigorous,
  qadir2015applying}. Yet many frameworks (e.g., P4~\cite{bosshart2014p4},
DPDK~\cite{intel2014data}, PF\_RING~\cite{pfring},
NetMap~\cite{rizzo2012revisiting}) and standalone implementations (Open
vSwitch~\cite{pfaff2009extending}, NetFilter~\cite{boye2013netfilter}) use
C. % These approaches require developers to drastically change or augment the way in which they write these network apps: massively annotate their code with logical derivations, use ``pure'' (side-effect-free) functions, abandon manual memory management and layout, and/or learn a new language.

Developers do not want to change the way they write code, they want to be able
to convince their users that the code they wrote is carrier-grade (and, if they
cannot prove correctness, to then fix the corner cases that impede the proof).
We conducted a small survey~\cite{verifsurvey} among NF developers from both
industry and academia, and almost 70\% insist on continuing to use
C. % instead of a higher level alternative. With current common practices and languages, it is not possible to achieve the desired level of confidence. The continued perception of unreliability and insecurity hinders adoption and therefore slows down progress (can we quote an operator who doesn't want to deploy stuff out of fear?).

The goal of our work is to enable developers to provide carrier-grade network
functionality in software without having to do a forklift upgrade of their
programming practices. In this paper we present a technique and a tool suite (a)
for formally verifying crash- and leak-freedom as well as the absence of several
classes of security vulnerabilities in NFs written in C/C++ (b) in a way that is
practical both in terms of developer effort and the performance of the resulting
software.

There exist approaches for providing formal guarantees for {\em stateless}
network packet processing code~\cite{dobrescu2014software}; to the best of our
knowledge, we are the first to provide such guarantees for real-world {\em
  stateful} software NFs without impacting performance. There are many such NFs:
all dynamic routing protocols (RIP, BGP, STP), NAT, BRAS~\cite{wiki:bras}, CDN,
IDS, DPI boxes, etc.

Practical verification of stateful software NFs is a difficult research problem.
Hence none of the verification techniques proposed so far can scale anywhere
close to verifying real-world stateful NFs. For the purpose of this paper, we
take as an objective to develop a NAT~\cite{rfc3022} box that is proven to not
have certain classes of vulnerabilities or crash or hang under any input. Even
though NAT is a classic network function, it has proven hard to get right over
time: the NAT on various Cisco devices can be crashed~\cite{cve-2015-6271} or
hung~\cite{cve-2013-1138} by carefully crafted inputs, with the same type of
problems existing in Juniper's NAT~ \cite{cve-2014-3817}, NATs based on
NetFilter~\cite{cve-2014-9715}, and the NAT in Windows Server~\cite{ms13-064}.
Thus, a NAT function that is formally verified to not crash, while providing
production-class performance, is an important milestone in this line of work.

We achieve both scalable verification and minimal developer effort by leveraging two insights into the verification of data plane software:
\begin{itemize}
\item {\bf Canonical data structures}: Unlike general-purpose software, data
  plane software employs a small number of commonly used network data structures
  (hash table, least-prefix-match, rings, etc.) and components. These lend
  themselves naturally to being abstracted into a library. They can be formally
  verified a priori, and their proofs then automatically leveraged for proving
  properties of the code that uses them.
\item {\bf Hybrid verification}: While typically one picks a verification
  technique and applies it throughout, no {\em single} technique works well for
  {\em all} of the code. However, data plane software offers us the opportunity
  to advantageously employ symbolic execution for high-level code (which is
  mostly stateless and has short execution paths) and verification based on
  separation logic to the library of components/data structures (which is highly
  stateful).

\end{itemize}

In this paper we present work that builds upon these insights and makes three
contributions:
\begin{itemize}
\item {\bf Speculative data plane verification} (SDPV), a technique for
  verifying stateful software NFs by fusing powerful formal verification with
  automatic symbolic
  execution. % It uses formal verification for data structures. It replaces them with simple models that encode some assumptions about the behaviour of the data structures. It then runs SE on the application code and later validates the used models. This approach postpones high-effort/low-return work and discards it when it turns out to be irrelevant.
\item {\bf The Vigor library of verified components} along with a tool chain
  that makes SDPV available to those who write software NFs in C/C++. The
  library consists of data structures (Hash Map, Batcher, Static Array) and a
  Port Allocator that are common for software
  NF. % The tool-chain brings together existing SE and formal verification engines to test the proposed technique.
\item {\bf The \vignat verified NAT box} that offers both high performance and
  is formally proven to not crash or leak memory. We estimate that it would take
  an NF developer about 1 week to develop and verify a NAT using Vigor.
\end{itemize}

In the rest of the paper, we give a short overview of Vigor
(\S\ref{sec:solution-overview}), describe our design (\S\ref{sec:design}), and
present our implementation (\S\ref{sec:implementation}). We then evaluate the
technique for two NAT prototypes (\S\ref{sec:evaluation}), discuss limitations
(\S\ref{sec:limitations}), present related work (\S\ref{sec:related-work}), and
conclude (\S\ref{sec:conclusion}).

%==============================================

\section{Solution Overview}
\label{sec:solution-overview}

\subsection{Developer experience}

This section motivates and explains at a high level what it's like to use Vigor
when developing NFs. For the sake of illustration, say we wish to develop a
trivial stateful NF that implements the discard protocol~\cite{rfc863}: an
infinite loop receives packets from one interface, discards the ones that target
port 9 (``discard port''), and forwards the others through another interface.
A possible implementation is on the Fig.~\ref{lst:orig}.

\begin{figure}[h!]
\begin{lstlisting}
#define CAP 512

int main() {
  struct packet packets[CAP] = {}; ^\label{line:pkt-def}^
  int begin = 0, queue_len = 0; ^\label{line:iter-def}^

  while(1) {^\label{line:loop-begin}^
    if (queue_len < CAP) {
      int end = (begin + queue_len)%CAP;
      if (receive_packet(&packets[end]) && ^\label{line:recv-pkt}^
          packets[end].port != 9)
        queue_len += 1;
    }
    if (0 < queue_len && can_send_packet()) { ^\label{line:can-send}^
      send_packet(&packets[begin]); ^\label{line:send-pkt}^
      begin = (begin + 1)%CAP;
    }
  }^\label{line:loop-end}^
  return 0;
}
\end{lstlisting}
  \caption{Original implementation of the Drop protocol. Contains an infinite
    loop (l.\ref{line:loop-begin}-\ref{line:loop-end}) and a latent ring
    implementation (declared on l.\ref{line:pkt-def},\ref{line:iter-def})}
  \label{lst:orig}
\end{figure}

In order to absorb bursts, the code employs a ring buffer. The application
interacts with the network using the following 3 functions:
\code{receive_packet}(l.\ref{line:recv-pkt}), which non-blockingly reads an
inbound packet and stores it at \code{packets[end]}, returning success or
failure; \code{can_send_packet}(l.\ref{line:can-send}), which checks if a new
packet can be sent; and \code{send_packet}(l.\ref{line:send-pkt}), which sends
the packet stored at \code{packets[begin]}. The application follows a common
practice: it consists of a main event loop, it has no other complex loops, and
it uses a ring, a popular data structure in network programming.

The ring keeps the state of our NF and makes many existing verification
approaches impractical.

Fig.~\ref{lst:vigor} shows the actual code that we are going to verify. The
first two steps show how we get there. \code{VIGOR_CHECK}
(l.~\ref{line:vigcheck}) adds additional property that we would like to verify
--- that we do not forward packets with port 9.

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);^\label{line:mk-ring}^
  if (!r) return 1;
  while(LOOP(1))^\label{line:loop-hd}^
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);
      VIGOR_CHECK(p.port != 9);^\label{line:vigcheck}^
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol.
    \code{LOOP}(l.\ref{line:loop-hd}) labels the infinite loop head and we
    explicitly use a ring(l.\ref{line:mk-ring}) from the Vigor library. \code{VIGOR_CHECK} at the
    line \ref{line:vigcheck} adds a semantic property to be verified.}
  \label{lst:vigor}
\end{figure}

{\bf Step 1: Label infinite loops.} Our approach requires special handling of
infinite loops in the application code. A developer puts a label \code{LOOP}
(l.\label{line:loophd}) into the condition in order to designate the
potentially infinite loop. For a typical NF there should be only a handful or
them (1 in our example).

{\bf Step 2: Use library components. } A developer replaces every array with a
component from Vigor library that performs the same function. For our example,
the library provides a packet ring with a simple interface: \code{bool
  ring_empty(ring*)} checks whether there is a packet in the ring, \code{bool
  ring_full(ring*)} checks whether the ring has run out of space, \code{void
  ring_push_back(ring*, packet* src)} adds packet \code{src} to the ring, and
\code{void ring_pop_front(ring*, packet* dst)} removes a packet from the ring
and places it in \code{dst}.

{\bf Step 3: Provide invariants.} Besides usual parameters, like capacity or
element size, most Vigor components also require simple logical invariants of
their elements to enable the reasoning about the application code. For our
example, the essential property of the ring is, that it contains only packets
with a port different from 9. A developer communicates that as shown in
Fig.~\ref{lst:customizations}.

\begin{figure}
\begin{lstlisting}
#define CAP 512
#define RING_CAPACITY CAP

/*@^\label{line:vf-user-prop-from}^
  fixpoint bool
  user_property(packet p) {
    switch(p) { case packet(port):
                  return port != 9; }
  }
  @*/^\label{line:vf-user-prop-to}^

static bool^\label{line:c-user-prop-from}^
packet_constraints(struct packet* p) {
  return p->port != 9;
}^\label{line:c-user-prop-to}^
\end{lstlisting}
  \caption{Custom definitions capturing the desired property. They capture the
    essence of every packet in the ring. Currently, we have to duplicate the
    packet condition in two languages (formal logic --
    ll\ref{line:vf-user-prop-from}-\ref{line:vf-user-prop-to}, and C --
    ll\ref{line:c-user-prop-from}-\ref{line:c-user-prop-to}) due to the
    limitations of our prototype.}
  \label{lst:customizations}
\end{figure}

{\bf Step 4: Run Vigor checker.} Vigor will find the appropriate symbolic model
and check that the application does not crash in any execution scenario. It will
then make sure that the chosen model is faithful to the formally proven
implementation of the component. \S\ref{sec:under-the-hood} provides more
details on how it is done.

{\bf Step 5: Link the NF code with the verified library} of components. The
resulting executable is guaranteed to be free of crashes, and hold the
properties formulated in \code{VIGOR_CHECK}, subject to the assumptions listed
in~\S\ref{sec:limitations}.

\subsection{Under the hood}
\label{sec:under-the-hood}

To verify that this NF does not crash, one could use the approach of Dobrescu et
al.~\cite{dobrescu2014software} and run exhaustive symbolic execution. The
symbolic execution (SE) technique entails using a symbolic execution engine
(SEE) to explore all paths through the code and check whether any failure
conditions hold --- in our case crash conditions. The inputs to the program are
marked ``symbolic,'' meaning that they subsume all possible input values, and
then the SEE simulates the execution of the program while propagating the
symbolic inputs along the dataflow paths into the program's variables. When
facing a branch (e.g., an \code{if} or a \code{while} loop), the SEE clones the
program state and takes each feasible path separately. Branches impose
constraints on variable values (e.g., an \code{if} statement imposes
restrictions on the values of the condition variable along the \code{else}
branch); the SEE collects these constraints along each path. When evaluating a
branch condition, the SEE asks an SMT constraint solver which of the branches
are feasible, given the constraints collected so far on that path prefix.

One could use an SEE like KLEE~\cite{cadar2008klee} to symbolically execute the
original application code (Fig.~\ref{lst:orig}) exhaustively --- once all paths
are found to be crash-free, the crash-freedom proof is complete. To perform SE,
it is necessary to provide symbolic models for the three network calls.
Something simple suffices, such as \code{receive_packet} generating a new
symbolic packet in \code{dst}; \code{can_send_packet} forking program state into
two states for the interface being busy and idle correspondingly; and
\code{send_packet} doing nothing.

Alas, naive symbolic execution would never terminate. The reason is ``path
explosion'': there is a large (often infinite) number of paths to explore, so
the SEE cannot finish. Path explosion results from cyclomatic
complexity~\cite{mccabe:cyclomatic}: every branch condition doubles the number
of paths, and every loop multiplies it by a factor; in our case, the infinite
loop gives rise to an infinite number of paths.

Vigor transforms infinite loops using loop
invariants~\cite[\S~2.1]{cormen2009introduction}. We added to KLEE support for
automatic havoc-ing~\cite{barnett2005boogie}. Havoc-ing simply replaces all the
variables that may change in the loop by unconstrained symbols at the beginning
of the iteration. In our example that would be the variables \code{begin} and
\code{queue_len} and the contents of \code{packets[0..CAP]}, but not the value
of the \code{packets} pointer because it never changes. The step 1 enables
Vigor to reduce each problematic loop to a single generalized iteration that
allows reasoning in one pass.

This change makes SE feasible, albeit impractical: for \code{CAP=5}, it takes 10
sec to exhaustively SE the code; for \code{CAP=30} it takes 15 min, for
\code{CAP=60} it takes 3 hours, and so on. Another classic SE challenge ---
``symbolic
indexing''~\cite{sen2005cute,godefroid2008automated,boonstoppel2008rwset} spoils
the party here. When
indexing an array with a symbolic variable (or when dereferencing a symbolic
pointer), the constraints on program state become extremely complex, because the
symbolic pointer or index could be referencing many different possible
locations. The complexity of the constraints chokes the SMT solver, and the SEE
grinds to a halt. In general, SE doesn't work for code with symbolic pointers or
offsets. Alas, stateful C code almost always has some of these.

In this example the symbolic indexing comes from the ring array. Once we
replaced the infinite loop with a single iteration we have to assume (remember
havoc-ing) the \code{begin} and \code{begin + queue_len} indexes may point
anywhere in the array.

In Vigor, we circumvent the problem of symbolic indexing by encapsulating program
state into well-defined data structures and putting them into a ``library of
components'' (step 2). This works well for NF development, because the NF code
uses a relatively small set of standard structures, such as ring buffers, hash
tables, least-prefix-match tables, etc. We provide to the SEE symbolic models of
these abstracted components that ``hide'' symbolic indexes and pointers from the
SEE.

After the step 2 exhaustive SE takes less than 100 millisec regardless of the
value of \code{CAP}. The fact that the code does not
crash on any path amounts to a proof of crash freedom. Or does it? The validity
of the proof depends on our library implementation of \code{ring} not
crashing; and (2) the symbolic model of \code{ring} accurately capturing how the
\code{ring} implementation works.

The Vigor library comes with a formal specification and a formal proof that the
implementation satisfies that specification. In other words, there is a formal
spec for \code{ring} and a proof that the \code{ring} implementation satisfies
the spec. This is how we enable SE to verify the high level code, while using
separation logic to verify the library of components --- two different
verifications techniques for two different parts of the complete application.
Doing the library verification requires deep expertise in program verification,
but our effort is amortized across all NFs that use the library.

Vigor makes it possible for NF developers to verify their code without having to
understand verification. Note that stating invariants(step 3) for dataplane
software turns out to require no verification expertise.

Here is an example of the proven specification for the function
\code{ring_pop_front}:

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p)
/*@ requires ringp(r, ?user_property,^\label{line:precondition-from}^
                   ?lst, ?cap) &*&
             lst != nil &*& packetp(p, _); @*/^\label{line:precondition-to}^
/*@ ensures ringp(r, user_property,^\label{line:postcondition-from}^
                  tail(lst), cap) &*&
            packetp(p, head(lst)) &*&
            true == user_property(head(lst)); @*/^\label{line:postcondition-to}^
{
  //@ extract_first(r); ^\label{line:lemma1}^
  struct packet* src_pkt = r->array + r->begin;
  p->port = src_pkt->port;
  r->len = r->len - 1;
  r->begin = r->begin + 1;
  if (r->cap <= r->begin) {
    r->begin = 0;
    //@ stitch_with_empty_overflow(r);^\label{line:lemma2}^
  } else {
    //@ stitch_with_empty(r);^\label{line:lemma3}^
  }
}
\end{lstlisting}
  \caption{A formal contract and verified implementation of the
    \code{ring_pop_front} function. The comments starting with \code{@} symbol
    are formal annotations that hold the function contract
    (\code{requires}(ll.\ref{line:precondition-from}-\ref{line:precondition-to})
    - precondition,
    \code{ensures}(ll.\ref{line:postcondition-from}-\ref{line:postcondition-to})
    - postcondition) and
    lemmas(ll.\ref{line:lemma1},\ref{line:lemma2},\ref{line:lemma3}) that help
    the proof checker to verify the state transformations in the code.}
  \label{lst:contract}
\end{figure}

This C code is intermingled with the correctness proof annotations that help the
formal verification engine (VeriFast) with small steps to transform the initial
state, given in \code{requires} to the final state, specified in \code{ensures}.
After proving the correctness of this transformation, the formal verification
engine will use \code{requires/ensures} formulas as pre- and post- conditions
for \code{ring_pop_front} contract. The trigraph \code{&*&} separates different
statements that hold together. For example

\begin{lstlisting}
ringp(r, user_property, tail(lst), cap) &*&
packetp(p, head(lst)) &*&
true == user_property(head(lst));
\end{lstlisting}
means:
\begin{itemize}
\item \code{r} points to a healthy ring of capacity \code{cap}, that contains
  \code{tail(lst)} packets, that satisfy the \code{user_property}.
\item \code{p} points to a packet ``\code{head(lst)}''.
\item ``\code{head(lst)}'' satisfies the \code{user_property}.
\end{itemize}

Note that \code{user_property(head(lst));} in the post condition ensures
\code{p->port != 9} property (Fig.~\ref{lst:customizations}).

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  ASSUME(packet_constraints(p));
}
\end{lstlisting}
  \caption{A symbolic model for \code{ring_pop_front}. It fills symbolically the
    a packet and adds a constraint to it.}
  \label{lst:symbolic-model}
\end{figure}

Some symbolic models work well for some code, while others work well for other
code, so the Vigor library comes with {\em multiple} simple models for each
component, and Vigor automatically chooses combinations of models. Once it finds
a combination of models with which the desired property (crash freedom, no
memory leaks, absence of certain security vulnerabilities) can be proven, Vigor
validates that the behaviors exposed by the model during verification are
consistent with the component's formal specification. Details on how it does
this appear in \S\ref{sec:design}.

A model of \code{ring_pop_front} used for this example is shown on
Fig.~\ref{lst:symbolic-model}. This model is simple enough to enable
exhaustive SE, but preserves the vital property --- \code{packet_constraints}
(Fig.\ref{lst:customizations}).

To validate the model fit for the particular use case, SEE dumps all the call
sequences (30 in this example) extracted from the execution paths. Vigor
checks whether the symbolic model has enumerated all the possible outputs for
each call sequence.

You can see one of the generated call sequences on Fig.~\ref{lst:call-sequence}.
It starts with \code{loop_invariant_produce} that marks the beginning of a loop
iteration. It continues then with calls to \code{ring_full} and \code{ring_empty},
noting (with \code{assume}) that both return \code{false}. The seqeunce ends
with a call to \code{ring_pop_front}. The last line detects underapproximation
of the model output.

\begin{figure}[h]
\begin{lstlisting}
struct ring* arg1;
struct packet arg2;

loop_invariant_produce(&(arg1));
/*@ open loop_invariant(_); @*/
bool ret1 = ring_full(arg1);
//@ assume(ret1 == false);
bool ret2 = ring_empty(arg1);
//@ assume(ret2 == false);
/*@ close packetp(&(arg2),
                  packet((&(arg2))->port)); @*/

ring_pop_front(arg1, &(arg2));
//@ open packetp(&(arg2), _);

/*@ assert(arg2.port != 9); @*/
\end{lstlisting}
  \caption{A call sequence, produced by SEE. Vigor adds formal annotations to
    constrain the execution space.}
  \label{lst:call-sequence}
\end{figure}

Formal verification engine checks this sequence using the contracts like in the
Fig.~\ref{lst:contract}, and ensures the abscence of misuse e.g. call for
\code{ring_pop} on a possibly empty ring. It then checks that the model output
covers all the possible outputs specified in the contract for the tip function.
This happens in the last line on the listing. The assert contains the
path-constraint generated by the symbolic model. In this case the contract
(Fig.\ref{lst:contract}) guarantees that the popped packet satisfies
\code{user_property} that entails \code{arg2.port != 9}, hence the trace is
legitimate. (see~\S\ref{sec:model-validation-criteria}).

\subsection{Misfit symptoms}

Consider potential model mismatches. Suppose we had an overapproximate model ---
Fig.~\ref{lst:overapproximate}. Then the \code{VIGOR_CHECK}
(Fig.\ref{lst:vigor}) in the application code would fail.
\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
  // No constraints on the packet port here.
}
\end{lstlisting}
  \caption{An overapproximate ring model fragment. It does not add the
    constraint necessary for the code to verify.}
  \label{lst:overapproximate}
\end{figure}

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  p->port = 0;
}
\end{lstlisting}
  \caption{An underapproximate ring model fragment. Instead of returning a
    constrained symbolic packet, it simply returns some particular packet, which
    clearly can not capture what the real ring would return on its place.}
  \label{lst:underapproximate}
\end{figure}

Suppose now we had an underapproximate model: Fig.~\ref{lst:underapproximate}.
Then the last line in the call sequence check (Fig.\ref{lst:call-sequence}) would look like:
\begin{lstlisting}
/*@ assert(arg2.port == 0); @*/
\end{lstlisting}
The assertion fails, because the contract (Fig.~\ref{lst:contract}) specifies
a wider range for \code{arg2.port} than just \code{0}.

% Note, that the \code{ring_pop} model does not even check whether it is called
% with an empty ring or not. The example application makes sure to call pop only
% if the ring is not empty. Therefore the check is not necessary. This model
% clearly overapproximates a ``ring'', e.g. \code{ring_empty} may return true
% even right after \code{ring_push}. This means, it may have caused false
% positives, but as long as it does not, no need to complicate it. For the sake
% of the example conciseness, we hard coded the invariant ``every packet has
% port != 9'' and the packet structure into the ring component itself. In a real
% component this detail should be abstracted and supplied as a parameter.

% Now a SE can analyze the code in an instant. As you can see the model does not
% maintain any state. It only \code{ASSERT}s that any pushed packet satisfy the
% invariant, and \code{ASSUME}s that every popped packet does. Last steps:

%========================================
\section{Design}
\label{sec:design}

In this section we describe the design of the Vigor technique and system.
Figure~\ref{fig:arch} provides a high-level overview of the three main
components of the tool chain. In order to prove properties of a network function
that uses the Vigor component library, Vigor symbolically execute the NF using a
symbolic execution engine (SEE); it then uses a Validator to combine the result
of the SEE with the formal spec and proof of the component library.

\begin{figure}[h!]
    \includegraphics[width=\columnwidth]{figures/impl_overview.png}
    \caption{The Vigor tool chain for automatic verification.}
    \label{fig:arch}
\end{figure}

%---------------
\subsection{Speculative Verification}
\label{sec:our-approach}

Vigor views a network function as a combination of stateless application logic
tied to components and data structures (collectively called ``components'') that
are stateful, such as hash tables, arrays, least-prefix-match tables, etc. In
the example of \S\ref{sec:solution-overview}, we encapsulated the \code{packets}
array into the \code{ring} component. Singular variables of course need not be
abstracted away, just large or variable-sized structures potentially accessed
with symbolic indexes. The role of the library is to group stateful components
into a unit that can be formally verified {\em once} an then {\em reused} across
all NFs---we can view the Vigor library as a domain-specific extension of the C
language. This makes sense because there is a relatively low number of
components needed to support a wide range of NFs. Figure~ \ref{fig:algo}
summarizes how Vigor works.

\begin{figure}[]
    \includegraphics[width=\columnwidth]{figures/vigor_algorithm.png}
    \caption{The Vigor verification workflow. Rectangles represent data, hexagons are the Vigor tools, and arrows indicate the flow of data.}
    \label{fig:algo}
\end{figure}

It is feasible for NF developers to develop verified components and add them to
the library. However, this requires a fair amount of verification expertise:
formally specify the component interface, implement it in a way that can be
verified but also has good performance, then produce a mechanically checkable
proof, write one or more symbolic models for it, and finally a plug-in for the
Validator. Vigor therefore adopts a usage model in which the vast majority of NF
developers just use the components in the Vigor library and run Vigor to prove
(or debug) their code.

% We had to find our own way to support high level data structures in SE,
% because the popular approach does not suit our goal. Many systems based on SE
% extend the target language with specification syntax allowing developers to
% formulate complex conditions. For instance,
% Frama-C/Value~\cite{kirchner2015frama} and VeriFast~\cite{jacobs2008verifast}
% use their own abstract specification language. We strive to make the
% verification simple and automated, but unfortunately current state of the art
% systems cannot reason automatically in such sophisticated terms and require
% manual small-step derivations. Developers end up spending more effort on the
% correctness proof than on the code itself.

%We explore an alternative way:

For each component of the library, Vigor provides several symbolic models. A
model is a concise over-approximating abstraction of how the specific component
behaves when invoked by NF code. A good model has very little state. When
symbolically executing an NF, the SEE uses one model for each component, and
explores all execution paths through the NF together with the models. If every
execution path has the desired property, then the NF has that property at all
times. Since models are relatively easy to write, it is conceivable for NF
developers to also write symbolic models.

A desired property is formulated as an assert. ``No division by zero'' is stated
as an assert that the denominator is non-zero; ``safe memory access'' is stated
as an assert that accessed memory is ``owned'' by the accessor. More details
appear in \S\ref{sec:properties}. If a symbolically executed code path completes
without any assert failing, we say the execution path was successful.

For each execution path, the Vigor SEE produces a trace of how the NF code
interacted with the components it uses (we define these traces formally in
\S\ref{sec:model-validation-criteria}). This is {\em speculative} in that we
don't have proof that the models are accurate abstractions of the components.
The question the Validator answers is whether these traces are consistent with
the formal specs of the components, i.e., could the interaction the NF code had
with the model have been a legitimate interaction with an implementation of the
component's specification. This is a weaker property than the model being a
correct abstraction of the component, but it is sufficient for our purposes.

Of course, execution paths may not always succeed during SE. While this may be
indicative of a bug in the NF, it could also be simply because the model is too
approximate. For example, in \S\ref{sec:solution-overview} a model of
\code{ring_empty} may be as simple as randomly returning true or false---for
some NF code, this behavior could crash it, if it assumes \code{ring_empty} to
return false right after pushing an element. A different model, however, keeps a
counter of how many elements are in the ring, and makes \code{ring_empty} return
true only when this counter is zero. The above NF code would not crash with this
model; on the downside, the more stateful model makes SE take longer. Therefore,
Vigor starts with the least-state models, and tries alternates until it either
finds a model with which SE succeeds on all paths or not.

If Vigor does not find any symbolic model for which all execution paths through
the NF code succeed, then it points the developers to the places where the
properties fail to hold, and helps the debug.

% \item The application developer provides loop invariants for long/unbounded
%   loops (e.g. event loop in the example). The invariants are mostly a
%   conjunction of the corresponding invariants for the symbolic models, but
%   they are likely to have an application specific part.

%   a number of speculative assumptions that we believe will hold for the real
%   implementations. In the example above some of these assumptions are: ``the
%   \code{ring_empty} function may return either 0 or 1'', ``if the
%   \code{ring_pop_front} returns success the popped packet’s port will never be
%   equal to 9'', and ``\code{ring_push_back} never crashes''.

% ^^^ do not forget to uncomment the references together with the figure
% \begin{figure}[h]
%     \centering
%     \def\svgwidth{\columnwidth}
%     \input{pie.pdf_tex}
%     \caption{Our conceptual division of the code into stateless and stateful parts with a strict interface between them.}
%     \label{fig:pie}
% \end{figure}

%-----------------
\subsection{Assumptions}
\label{sec:assumptions}

Our approach makes several assumptions. First, we assume that our symbolic
models are over-approximations of the environment in which the NF runs (network
stack, OS, etc.). In devising these models, we aim to include every possible
behavior, but the absence of a formal specification of the environment makes it
impossible to validate the models. Second, our guarantees rely on the hardware
running the software dataplane correctly. Third, we assume the compiler
implements the same language semantics employed by our tools (e.g., same byte
length for C primitive types).

Thus, the trusted computing base for a Vigor-verified NF consists of the
compiler, the Vigor tool chain (including its formal verification engine and
SEE), the network stack, device drivers, OS kernel, BIOS, and hardware.

% We distinguish of the verified component library from development of an
% application using this library. These two activities require different skills
% and ideally should be done by different people.

% This paper proposes the following approach expressed in a number of steps. We
% use three tools: ``SE verifier'' stands for our SEE, ``formal verifier'' is
% our formal verification proof checker and assistant and a ``validator'' ---
% the tool that validates legitimacy and exhaustiveness of the model behaviour
% post factum. We differentiate human work between two developers: a component
% developer and an application developer. During the development of our NAT box
% they were the same person.

%------------------------------
\subsection{Handling pointers}
\label{sec:handling-pointers}

Pointers are the soul of the C language---they embody the power and the danger
of low-level systems code. A NF focused on performance can not avoid them. We
embrace them and support the most used idioms in our tool.

An important choice is the memory allocation strategy. Dynamic allocation offers
more flexibility but incurs performance penalties. Static allocation should be
preferred because it allows better control over layout, making it possible to
optimize for cache and same the allocation overhead at runtime. However static
allocation reduces flexibility, e.g. it fixes the component capacity at the
compile time.

Among all the pointers appearing in the execution context, only the pointers
crossing API boundary are important for the model validation. We divide these
pointer into two kinds: argument pointers and return pointers.

An argument pointer may be either:
\begin{itemize}
\item The address of the component state structure. It is an equivalent of
  ``this'' pointer for C++/Java or ``self'' reference for Rust/Python. We do not
  inspect the memory by such pointers, we only note the aliasing between them.
  The component owns this memory and only the formally verified code can access
  it.
\item Output parameter. The application owns this memory. Therefore we have to
  trace the evolution of the pointee value. The pointee value before the call,
  which we trace as the function input, and the pointee value after the call,
  which goes to the function output. A special case of output pointer is a
  double pointer to the component state structure (X** p), set by the component
  allocation function. The app then owns the pointee (*p), and the SE verifier
  traces it, but the pointee of the pointee (**p) is the component internal
  state, so no need to trace it. We currently do not support other cases of
  double or deeper pointers (e.g. an output parameter to be set to point to a
  piece of the internal state).
\end{itemize}

A return pointer for our components always points to a piece of the internal
state (e.g. an array cell). We trace the pointee and also make sure the
application does not access it when the component regains the ownership.

%--------------------
\subsection{Property Formulation}
\label{sec:properties}

Vigor can verify any property that can be stated as a safety property, i.e.,
formulated as an \code{assert} condition. This includes various sources of
crashes (dereferencing a null pointer, reading from an unmapped memory page,
dividing by zero, etc.) as well as several classes of security bugs related to
memory safety (writing to unallocated memory, writing to the wrong memory
object, use-after-free, double-free, etc.) and related to integer operations
(e.g., integer overflow).

When symbolically executing a code path, the SEE checks whether any assert fail.

To check for memory leaks (or leaked resources in general) is a bit more
complicated, because that is a global property. Vigor therefore must keep track
of memory ownership. In the application logic we forbid dynamic memory
allocation, and all allocated memory is controlled in the library, thereby
giving the formal verifier explicit visibility into memory ownership. The
temporary pointer access control enables the SE verifier to ensure the
application returns all the memory chunks taken from a component and never
accesses them after that.

Vigor does not specifically deal with semantic properties (e.g., dropping a
packet, using an incorrect MAC address as a destination), but many of these
properties can be formulated by NF developers as \code{assert} statements in
their code. Vigor will then determine whether those \code{assert}s hold on all
paths.

%----------------------
\subsection{Validating a Symbolic Model}
\label{sec:model-validation-criteria}

To validate a model behavior under certain circumstances means to ensure the
behavior agrees with the specification, or is more relaxed (i.e., admits more
behaviors than the spec). Suppose that, for the input \(I\) to the prefix of an
execution trace (a ``run''), the specification produces the output $O_s$, and
the model produces the output \(O_m\). Then the model is valid iff for all call
sequences performed during the exhaustive SE, \(O_m\) contains \(O_s\).

More formally:
\begin{enumerate}
\item \(F^i\): The function input (\(F^i = fun\_name + args +
  arg\_pointee\_vals\_before + constraints\)) consists of the function arguments
  (possibly symbolic), and constraints bounding these arguments (specifically,
  the transitive closure of all the constraints involving the symbols mentioned
  in the formulas for the arguments). For a pointer argument (which must point
  to a concrete address --- a practical limitation), the pointee must be
  considered according to the \S\ref{sec:handling-pointers}. We use the
  numerical value of the pointer to detect aliasing between the variables, and
  then throw it away, since we forbid pointer comparison anyway.
\item \(F^o\): The function output (\(F^o = ret\_val + pointee\_vals\_after +
  constraints\)) is its return value and the pointee values for the output
  pointer arguments, as well as the constraints for all the symbols involved
  (transitive closure, same as for \(F^i\)).
\item \(I\): The run input (\(I = <F^i,F^o>_{i=0..n-1} + F^i_n\)) is a set of:
  function names, function inputs(\(F^i\)) for all the functions in the sequence
  and function outputs(\(F^o\)) for all but the last call.
\item \(O\): The run output (\(O = F^o_n\)) is the output of the last API
  function call in the sequence.
\item \(C\): A call sequence(= run), (\(C=I+O\)) consists of a run input and
  run output. It summarizes a particular interaction scenario with the model. It
  consists of the calls to component API functions (with function names,
  arguments and return values), and leaves out all other details like other
  calls or unrelated path conditions.
\end{enumerate}

Let us represent the formal specification and the symbolic model total states by
large sets. A total state consists of variables V, bound by logical formulas F.
So, the specification and model total states are:
\[
  \begin{aligned}
  &\mathrm{total~specification~state} = \{V_s : F_s(V_s)\}\\
  &\mathrm{total~model~state} = \{V_m : F_m(V_m)\}
  \end{aligned}
\]
The output is then simply a projection of the corresponding state. The
projection (\(P\)) gives the output variables (return value and output pointers of
the tip call):
\[
OV_s = P_s(V_s): OV_m = P_m(V_m).
\]

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/projections.png}
    \caption{The outputs \(O_m, O_s\) represented as projections of the total
      formal specification and model states.}
    \label{projections}
\end{figure}

With the introduced definitions, the following formulas define the outputs:
\[
  \begin{aligned}
    &O_s = \{P_s(V_s): \forall V_s F_s(V_s)\} \\
    &O_m = \{P_m(V_m) : \forall V_m F_m(V_m)\}
  \end{aligned}
\]
Then, the model exhibits all the specified behaviours (and may be some more) iff
\(O_s \in O_m\), or:
\[
\forall V_s : F_s(V_s) \exists V_m : F_m (V_m) \cap P_s(V_s) = P_m(V_m)
\]
This formula in general is too difficult for the existing SMT solvers because of
the quantifier. However, due to similarities in the model and the specification
it is easy to find appropriate values for \(V_m\) --- they happen to be some
values from \(Vs\). Let us call this assignment \(A\). \(A\) by construction
satisfies the equality \(P_s(V_s)=P_m(A(V_s))\). So the condition becomes:
\[
\forall V_s : F_s(V_s) \rightarrow F_m(A(V_s))
\]
--- which we formulate as an assertion and let the formal verification engine
decide its validity.

This validation criteria guarantees that the formally verified implementation
will never do anything that the model did not explore during verification. So
each possible behaviour of the real execution (and may be some extra) was
checked during the SE, and it is 1) allowed by the specification and 2) does not
cause the application code to crash.

\subsection{Algorithmic Requirements}
\label{sec:algorithmic-requirements}

For a NF to be verifiable by Vigor, it must fulfill several requirements. These
are fully aligned with best practices in developing NFs.

First, the application must contain no deep recursion and only use well
structured loops. For long or unbounded loops, a developer must supply an
invariant. Most NFs naturally satisfy this requirement due to performance
considerations.

Second, program state must be encapsulated in a set of formally specified
components (e.g., port allocator) or data structures (e.g., hashtable). Vigor
uses the component API as a boundary between the code that is automatically
verified using SE and the code that is manually
verified.% (Figure~\ref{fig:pie}).

Third, there must be a well defined OS interface. As it is infeasible to
symbolically execute the whole OS, the SEE stops at the OS interface boundary.
Behind this interface we trust the correct implementation and we use symbolic
models of the OS API functions for SE. It is conceivable to use a verified OS
kernel, like SeL4~\cite{klein2009sel4}, and thus have a fully verified stack.

Finally, the NF must implement a classic event loop; most NFs already do so. For
Vigor specifically, there must be a single infinite loop, with each iteration
dedicated to processing a single event (an incoming packet). Within each
iteration, the sequence of interactions between the code and the library
components and the OS must be relatively short --- the complexity of models and
the amount of symbolic state are often proportional to the length of the API
call sequence it needs to support.

\subsection{Language Restrictions}
\label{sec:language-restrictions}

When writing NFs to be verified with Vigor, developers are free to use C.
However, we impose a few restrictions that ensure good isolation between
application code and the library components: We require that the
component-related memory be accessed only through the library API. There should
be no pointer comparisons except for equality --- the C99 standard anyway
prohibits most pointer comparisons, and we extend this restriction in order to
better isolate the component implementation from the code that uses it. Finally,
pointer accesses must be both domain-limited and time-bounded: certain pointers
may access only predefined range (i.e. a single unit of the declared type) and
only on a restricted segment of the execution, between the corresponding API
calls.

These restrictions do not apply to the formally verified component library.

% \subsection{Component Design Choice}

% A component designer faces a choice between conciseness and performance. Low
% expressive power of C forces the component developer to implement an
% abstraction either through typeless pointers or macros. Pointer implementation
% introduces an extra dereference, which may hurt performance. Macros are a zero
% cost abstraction. They perform better as they avoid extra pointer dereference
% and inline-friendly. Unfortunately, they hurt conciseness and readability of
% the code. In the long stretch a generics language extension such as C++99
% templates could alleviate this trade-off.

\section{Implementation}
\label{sec:implementation}

In this section we describe the prototypes that embody the ideas presented in
the previous section. We then use these prototypes to evaluate Vigor on the
development of a NAT middlebox, described in \S\ref{sec:evaluation}.

\subsection{Vigor Tool Chain}

To assess the practicality of Vigor technique, we implemented the three tools
described in the section~\ref{sec:our-approach}. The SEE is responsible for the
application verification, the Formal Verifier checks correctness of the
components’ implementation in the library, and the Validator ensures the two
verifiers agree on the component APIs used.

\paragraph{SE verifier} We built our verifier for speculative SE by modifying an
open source C/C++ SEE KLEE~\cite{cadar2008klee}. By default KLEE terminates
execution paths that are too long or take too much memory. We disabled this
feature to ensure it explores all the paths. We also extended KLEE with three
new features: Partial loop invariant support. The analysis finds automatically
the variables that may change in the loop, and havocs them. Currently, a
developer have to manually insert the assertions and assumptions according to
the loop invariant definition. Dynamic pointer access control. The patch
provides primitive intrinsics that allow a component developer to open and close
access to a pointer between the API calls. API call sequence dump. This patch
summarizes each control path in the program extracting the component API calls
and dumps them for the Validator.

\paragraph{Formal verifier} Our verifier is essentially
VeriFast~\cite{jacobs2008verifast} extended with the ability to export the
abstract state (\(O_s\)) for the Validator. Specifically we export the
accumulated conditions for the output variables at the end of the API calls
sequence.

\paragraph{Validator} We wrote this third tool completely from scratch.
Validator independently analyzes each API calls sequence. We exploit the data
parallelism using GNU parallel tool~\cite{Tange2011a}. For each API call
sequence in the KLEE dump, Validator first transforms it into VeriFast task. A
VeriFast task is essentially a C source file, containing the sequence of calls.
The Validator enriches the sequence with:
\begin{itemize}
    \item Symbolic variables used in the sequence.
    \item The constraints that form the sequence context.
    \item Lemmas to help formal verification.
\end{itemize}
Next, it runs VeriFast to check that the specification permits this particular
sequence of calls under the provided constraints. Finally, the Validator imports
the specification output state \(O_s\) and checks the inclusion of the
specification state \(O_s\) to the model state \(O_m\) using Z3~\cite{de2008z3}
SMT solver (see Section~\ref{sec:model-validation-criteria}).

Figure~\ref{fig:validator} shows the data flow between the Validator elements.

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/validator.png}
    \caption{The internal Validator dataflow.}
    \label{fig:validator}
\end{figure}

\subsection{Verified Component Library}

For our Vigor library prototype, we implemented four stateful components. These
are sufficient to develop the NAT middlebox of \S\ref{sec:evaluation}: A hash
map keeps current flows, and allows lookup by two keys corresponding to the
packets coming from internal and external interfaces; the Port allocator is
responsible for allocating new ports and expiring inactive entries; the Batcher
allows batch processing, keeping outbound packets until the NIC is ready to send
them; and the Static Array instances keep various configuration values. We are
careful in how we abstract components like the Static Array, because NFs are
expected to access them on hot paths, and so it is important that the data
reside in the processor cache.

\paragraph{Hash Map}

We implement a two-key hash map using three parts: integer indexed array of
flows and two single-key hash maps. The array stores flows together with the
keys. The hash maps internally are fixed size arrays of tuples $<$busy-bit,
key-hash, key-ptr, flow-index$>$:
\begin{itemize}
    \item Key-hash is the value of the hash function for the key.
    \item Key-ptr is the pointer to the full key for precise comparison.
    \item Value-index is the index in the array of values.
    \item Busy-bit flags the tuples that are not garbage.
\end{itemize}

Flow insertion linearly searches a free cell (with busy-bit = 0) from an index
equal to the hash of the corresponding key. To insert a flow, user must also
supply an unused index. We use port allocator to allocate these indexes.

Flow search and removal also starts from the hash-induced index looks for a cell
with the same hash, with confirmation by a total equality of the stored and
searched keys.

This implementation is not efficient for some cases. For instance, if the flow
is not present in the table, the algorithm will traverse the whole table and the
search time will be proportional to the capacity.

To address this inefficiency, we optimize the single-key maps. We add another
bit into the tuple. These bits group cells into hash-collision chains, and the
map searches for a flow only across that chain, not across the entire table.
Unfortunately, we did not yet verify the optimized implementation. We estimate
re-verification to take about half of the initial verification effort (see
section~\ref{sec:efficiency})

A hash map is a widely used data structure. It allows to store and retrieve
entries by arbitrary complex keys. Such functions as L2 switching, ACL
filtering, IPS, etc. all use hash tables.

\paragraph{Port Allocator}

Port allocator keeps two chains \emph{alloc} and \emph{free} packed into an
array. The size of the array equal to the port range available to allocator, and
each cell corresponds to a port to be allocated. \emph{Alloc} chain is a double
linked list. It contains all the allocated ports with last access timestamps
sorted by these timestamps. \emph{Free} chain is a single linked list. It keeps
track of the available ports.

All operations take constant time, and only manipulate the cells of the array.
Allocate operation moves head of the free chain to the head of the alloc chain.
Access moves the element from the middle of alloc chain to the head. Expire
takes the tail element from the alloc chain and moves it to the head of the free
chain.

The port allocator is essentially a number allocator from a predefined range.
Constant time allocation and expiration makes it a basic component for any
application that manages resources in run time, such as L2 switch or L3 router.

\paragraph{Batcher}

Batcher is an array with a counter. It allows to store elements one by one (up
to a preset capacity), and to release them all at once. Almost every DPDK
example uses a batcher. It is necessary to achieve high throughput, because NICs
work better on batches, rather than on individual packets.

\paragraph{Static Array}

Every DPDK example uses multiple static arrays for various configuration
parameters, counters and other statistics. Our static array is a thin wrapper
around a regular C array. This wrapper adds zero overhead, but enables our
tool-chain to control access to the array and to substitute the whole array with
just a couple of bytes during SE.

%===================

\section{Evaluation}

\label{sec:evaluation}

In this section we answer three questions:
\begin{enumerate}
\item Is our approach effective at verifying real-world stateful NFs? In
  \S\ref{sec:effectiveness} we provide a qualitative answer to this question for
  a DPDK-based NAT middlebox.
\item How efficient is our approach in terms of human and compute effort
  (\S\ref{sec:efficiency})?
\item Is the performance of the verified NAT box good enough for practical
  real-world use by sysadmins and network operators (\S\ref{sec:performance})?
  % \item What are the trade-offs inherent to Vigor, and how do they impact
  %   verifiability and performance? In \S\ref{sec:tradeoffs} we give the reader
  %   a deeper understanding on how the various parts of the technique
  %   contribute to its scalability and cost.
\end{enumerate}

As a proof of concept, we implemented and verified \vignat, a NAT middlebox
based on the DPDK framework~\cite{intel2014data}. We use NAT to refer
specifically to Dynamic Source Network Address-Port Translation~\cite{rfc3022}.
We reused boilerplate code from the DPDK L3 forwarding example. \vignat has
1,175 LOC, and it links with the Vigor library that has an additional 941 LOC.
\vignat supports TCP and UDP packet forwarding with dynamic source address--port
translation between multiple internal and one external interfaces. It supports
connection initiation from an internal interface and refuses connections from
the external one. It supports concurrent flows up to a configurable maximum
number; flows expire after a configurable period of inactivity.

% The biggest difference of our application code from a usual code you can see
% in network applications is array access. To get an element from our Static
% Array component, due to our memory ownership technique, a developer have to
% explicitly borrow it through API call and then return the borrowed element to
% access another one.

Our evaluation uses two versions of this NAT box. During our experiments we
discovered a silly performance bug in our hash table implementation that led to
excessive hash collisions under certain circumstances. We fixed it, but did not
complete the re-verification by the submission deadline. We therefore report
results for both \vignata (100\% verified but slow) and \vignatb (verified,
except for the hash table).

%---------------------------
\subsection{Vigor's Effectiveness}
\label{sec:effectiveness}

The ultimate goal of our work is to give operators the confidence that
software-based NFs will not undermine the robustness and security of their
network. It took about 12 minutes on a laptop for Vigor to prove that \vignat is
free of crashes, free of memory leaks, free of memory overflows, and free of
integer overflows. We verified \vignat only for one internal and one external
interface, but do not foresee any challenges in doing so for multiple
interfaces.

Beyond merely providing a binary yes/no answer to the question of whether an NF
has a given property or not, Vigor also helps developers find those corner cases
that impede the proof. There are three classes of bugs that Vigor uncovers in
the verification process, and as the citations show, these appeared in the
released Open vSwitch and DPDK code:
\begin{enumerate}

\item Security vulnerabilities~\cite{bug:ovs-uaf, bug:dpdk-uaf1, bug:dpdk-uaf2,
    bug:ovs-faf, bug:ovs-overflow, bug:dpdk-overflow1, bug:dpdk-overflow2,
    bug:ovs-int-overflow, bug:dpdk-int-overflow} in the form of memory overflows
  or integer overflows. These can enable an attacker to gain remote control of
  the middlebox.

\item Crashes~\cite{bug:ovs-0ptr, bug:dpdk-0ptr1, bug:dpdk-0ptr2,
    bug:dpdk-0ptr3} destabilize the network. Furthermore, many of these can be
  triggered by a remote attacker with a carefully crafted packet (e.g., trigger
  a bad pointer dereference or a division by zero), thereby opening up the
  middlebox to denial-of-service attacks.

\item Memory leaks~\cite{bug:ovs-mem-leak1, bug:ovs-mem-leak2,
    bug:ovs-mem-leak3, bug:ovs-mem-leak4, bug:dpdk-mem-leak1,
    bug:dpdk-mem-leak2} are often exploited for denial-of-service by exhausting
  the middlebox's memory. Vigor's memory ownership model makes them both easy to
  find and to debug.

\end{enumerate}

Vigor does not support today proving concurrency properties or higher level
semantic properties. Nevertheless, we consider the verifiable absence of
low-level flaws to be a solid step toward complete verification.

%---------------------------
\subsection{Efficiency of Using Vigor}
\label{sec:efficiency}

A key question when using Vigor is how much longer does it take a developer to
produce verified code vs. non-verified, and how much harder is it. Another
question is how much effort it takes to develop and verify the component
library, but that effort gets amortized across many uses and is therefore less
important.

\vignat's 1,175 LOC require about 30 lines of developer annotations (3\%
overhead) in the form of loop invariants and loop labels like \code{LOOP(1)} in
\S\ref{sec:solution-overview}). Static analysis can infer all the other
information needed for the proofs. In our current Vigor prototype we have not
implemented this static analysis, and so our \vignat prototype has 176 lines of
annotations (15\% overhead)in the form of C structure layout description,
conditional compilation constructs, etc. that make up for simple but
unimplemented analyses in Vigor.

Our estimate is that it would take an NF developer about 1 week to develop and
verify \vignat based on reading this paper (in particular
\S\ref{sec:algorithmic-requirements} and \S\ref{sec:language-restrictions}). Our
own development time is less representative, both because of our understanding
of the system and the fact that we co-developed Vigor with \vignat.

Once the code is developed and annotated, verification is fast. On an Ubuntu
16.04 Linux laptop with an Intel I7-6600U CPU and 16GB LPDDR3 RAM, exhaustive SE
of \vignat takes 2 minutes and 650 MB of RAM, and the Validator takes 10 minutes
to validate the 700 call sequences that result from SE. Verification of the
library's correctness proof takes 15 seconds.

\medskip

Developing and verifying the library took about 5 person-weeks, not counting our
learning curve and various setbacks. Detailed data appears in Appendix B.

Table~\ref{tab:loc} shows the size of each of the library's four components. As
expected, the formal specification and proof dwarf by a 10-20$\times$ margin the
implementation. This is why we consider Vigor's design---using SE for NF code on
the one hand, and heavy verification for the reusable library on the other
hand---to provide a favorable balance for real-world NF development. The
symbolic model and the validator plug-in are more concise than the component
implementation, which makes it feasible to provide several of them for each
component, to suit different usage patterns.

\begin{table*}[h!]
\centering
\caption{Size (in LOC) of the parts required to develop and prove each  library component and data structure.}
\label{tab:loc}
\begin{tabular}{lllllll}
\hline
  Component             & Interface & Implementation & Formal spec & Symbolic model & Proof & Validator plug-in \\ \hline
{\bf Hash Map}        & 29         & 688            & 596                  & 281            & 4,078         & 450              \\
{\bf Port Allocator} & 22         & 210            & 310                  & 95             & 4,439         & 129              \\
{\bf Static Array}   & 8          & 15             & 31                   & 145            & 155          & 39               \\
{\bf Batcher}        & 22         & 28             & 27                   & 60             & 65           & 23               \\ \hline
\end{tabular}
\end{table*}

%---------------------------
\subsection{\vignat Performance}
\label{sec:performance}

In this section we address the question of whether verifiability comes at the
price of performance. We evaluate \vignat's forwarding throughput and latency,
comparing it to a NAT based on NetFilter~\cite{boye2013netfilter}, which is
widespread and used in all the Linux-based home and small-enterprise routers. We
set up the NetFilter masquerade rules as shown in Appendix A, and we tune the
Linux kernel for optimal performance as described in~\cite{nf-testing}. We label
it {\em LinuxNAT} in our plots.

As mentioned earlier, we report on two versions of our NAT: \vignata has a
performance bug in the hash table, whereas \vignatb does not, but its hash table
is not yet verified. We configure both to have a maximum capacity of 61,000
flows. We use three identical machines, each a 32-core Intel Xeon E5-2667 v2 @
3.30GHz with 32GB RAM with i350 1 Gbps DPDK-compatible NICs.

% The throughput results are stable, so we do not collect statistical
% significance. However latency varies noticeably and we measure repeat the
% experiment ten times to reveal the variability.

%-------------------------
\paragraph{Throughput.}

We set up two nodes in the topology shown in Figure~\ref{fig:throughput-setup},
as per RFC 2647~\cite{rfc2544}: a tester machine connects to the middlebox over
two interfaces. We use \code{DPDK-Pktgen}, a high-performance traffic generator;
it sends packets through {\em outbound} and counts them on the {\em inbound}
interface. It assigns each packet a port from a given range to generate the
corresponding number of distinct flows. We run 40-second measurement intervals.

\begin{figure}[h!]
    \includegraphics[width=\columnwidth]{figures/throughput_setup.png}
    \caption{Network topology for throughput experiment.}
    \label{fig:throughput-setup}
\end{figure}

In Figure~\ref{fig:throughput-overall} we show the maximum packet rate achieved
with a maximum of 1\% packet loss, for different numbers of concurrent flows. At
the outset, \vignat outperforms LinuxNAT; this is in no way related to Vigor but
rather is due to DPDK and to LinuxNAT being interrupt
driven. % Our prototype is not optimized either, but the fact that it uses DPDK makes it much more efficient. %; compared to a trivial DPDK application that merely forwards packets, \vignat does not add

\begin{figure}[]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/brpt.pdf}
    \caption{Throughput as a function of the number of concurrent flows, with maximum 1\% loss rate.}
    \label{fig:throughput-overall}
\end{figure}

\vignata does well up until $\approx$28K connections, at which point it
collapses due to the performance bug. In contrast, \vignatb sticks to the
zero-loss line up until $\approx$59K connections, at which point our 61K-entry
hashtable experiences too many collisions to keep up---increasing the size of
the hash table would move this point further to the right.

To zoom in on what happens around those two inflection points, we plot two
throughput graphs. Figure~\ref{fig:throughput-27900} shows what happens for
27,800 concurrent flows: the hash collision bug in \vignata starts causing it to
grind to a halt at specific send rates; this behavior gets rapidly worse as the
number of concurrent flows increases. Figure~\ref{fig:throughput-59k} shows what
happens for 59,000 concurrent flows: \vignatb's flow table fills to 95\% and
thus slows down, starting to perform comparably to LinuxNAT on average, but with
significant jitter. As the number of flows increases, \vignatb slows down
further. % Reducing jitter is part of future work.

\begin{figure}[]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/rt27900.pdf}
    \caption{Packet rates for 27,900 concurrent flows.}
    \label{fig:throughput-27900}
\end{figure}

\begin{figure}[]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/rt59k.pdf}
    \caption{Packet rates for 59,000 concurrent flows.}
    \label{fig:throughput-59k}
\end{figure}

%---------------
\paragraph{Latency}

To assess \vignat's latency characteristics, we use the network topology in
Figure~\ref{fig:latency-setup}. We use separate ports on the tester for flow
generation and latency sampling. We use again DPDK's \code{PktGen} for packet
generation. We sample latency with netperf~\cite{jones1996netperf}
\code{TCP_RR}, which measures the request-response delay to \code{netserver} and
computes the average over 80 seconds.

\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/latency_setup.png}
    \caption{Network topology for latency measurements.}
    \label{fig:latency-setup}
\end{figure}

We also wrote a trivial DPDK application, derived from the DPDK l2fwd
example~\cite{dpdk-l2fwd} which only forwards packets. To eliminate the routing
overhead, we replaced L2 routing with a simple MAC address swap. This
application is stateless and performs only a couple writes to the packet header,
so we consider it the lowest possible latency one could achieve with DPDK; we
label it {\em Plain forwarding} in our plots.

Figure~\ref{fig:latency-plot} shows one-way latency as the number of concurrent
connections through the middlebox increases. The performance issue in \vignata
causes its latency to increase with the number of flows, but \vignatb stays
flat. There is increasing jitter as the number of flows increases, and
investigating this is part of future work (we believe it's due to DPDK). \vignat
has on average 23\% higher latency than LinuxNAT, but nevertheless stays under
111~$\mathrm{\mu sec}$, providing sufficient performance for many settings. Note
that DPDK is designed to sustain high throughput, so it uses batching
aggressively, which hurts latency---\vignat introduces a mere 5\% overhead
compared to {\em Plain forwarding}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/lat.pdf}
    \caption{Latency vs. number of concurrent flows.}
    \label{fig:latency-plot}
\end{figure}

%---------------
\bigskip

If the reader believes our assertion that we can verify \vignatb's hash table in
a few person-days, then the results in this section imply that developing a
verified NAT middlebox does not come at the expense of performance. NF developer
effort is relatively low, which makes this a practical approach. We therefore
conclude that Vigor can succeed in its objective of enabling developers to
productively write software-based data plane NFs that simultaneously meet the
highest standards of reliability, security, and performance.

\section{Limitations}

\label{sec:limitations}

Vigor's greatest limitation is that it is not fully automated yet. In
particular, we expect NF developers have to write loop invariants for long and
infinite loops. These loop invariants are essential to enabling scalable
exhaustive SE with our modified KLEE. As mentioned before, coming up with these
invariants for NF code is typically not difficult. Nevertheless, one of our
immediate future work items is to leverage invariant induction
techniques~\cite{perkins2004efficient, yang2004automatically} to automate at
least part of this task.

Application developers may also encounter ``false positives'', i.e., Vigor fails
to prove the desired property. As mentioned before, this can be due to the
property indeed not holding, or because the symbolic models are too
over-approximate. The NF developer can either develop their own more detailed
model, or request the component library developers to provide a suitable model.

Besides the trusted computing base described in \S\ref{sec:assumptions}, Vigor
relies on the NF developer to state correct invariants, as well as the library
developers to correctly annotate the symbolic models with KLEE intrinsics (so
that they produce correct traces). Finally, we trust the NF developer to not
link their application with unverified components. If any of these assumptions
fail to hold, Vigor cannot guarantee any properties.

With respect to our NAT use case, as already mentioned we have two versions, one
of which is not fully verified (but will be shortly). Also, our NAT does not
participate in any routing protocols; it needs an L3 router on each interface to
dispatch the packets to the rest of the network. We supply the MAC addresses of
these routers as a runtime argument to the NAT, and can not update it
on-the-fly.

\section{Related Work}

\label{sec:related-work}

%In this section we review some of the similar projects undertaken in the past.

Dobrescu et al.~\cite{dobrescu2014software} analyzed modular NFs. They managed
to analyze many stateless Click elements using exhaustive SE, and exploiting
modularity and low cyclomatic complexity of such applications. However, their
original approach is very limited in verification of stateful elements. In
addition to crash freedom they managed to verify bounded execution and filtering
properties. Doing this in Vigor for stateful NFs is future work.

SeL4~\cite{klein2009sel4}, CompCert~\cite{stewart2015compositional}, and
FSCQ~\cite{chen2015using} show how to do functional verification of practical
systems. All of them prove the semantics of the system. Unfortunately they all
require the use of a high-level programming language and deep expertise in the
verification domain, which is unacceptable for our problem.

Beringer et al~\cite{beringer2015verified} applied formal verification to an
OpenSSL implementation, proving its functional correctness and cryptographic
properties. They used certified tool chain, reducing trusted code base just to
just a small CoQ proof checker core (~8k LOC). While it paves the road for
formal verification towards real systems, it is not yet suitable as a common
development practice.

SymNet~\cite{stoenescu2013symnet} uses SE to reason about whole network
configuration with per-flow state. They do not verify the middle box
implementations, instead parse the configuration files and use special models in
place of the components. Their work naturally complements ours.

IronFleet~\cite{hawblitzel2015ironfleet} also aims for a verification of
applications distributed over a network. They use a different approach, though.
Habwlitzel et al develop IronFleet from scratch in Dafny --- a high level
language with built-in verification support.

Others checked existing distributed systems. Pedrosa~\cite{pedrosa2015analysis}
and Guerraoui~\cite{guerraoui2011model} explored the global state tree down to a
certain depth. Guerrauoi et al. used a similar approach to ours: they made
reasonable assumptions that simplify verification and validated the ones that
turned out important.

CEGAR~\cite{clarke2000counterexample} is similar to our technique in the spirit
of optimizing out the high effort/low return work. It only refines the models
when given a clear evidence of the their insufficiency. Clarke at al.
demonstrated a very generic approach. We focus on the networking domain, and
exploit the features inaccessible for CEGAR, like the low number of data
structures and simplicity of the application logic.

\section{Conclusion}

\label{sec:conclusion}

We presented a technique for proving properties of software NFs. We exploit the
specifics of NFs to combine the benefits of symbolic execution with formal
verification. Our experimental results demonstrate the practicality of the
approach for the development of NAT box that is guaranteed to have no memory
leaks, never crash, be free of certain security vulnerabilities, yet still
performs on par with NetFilter.

{\footnotesize \bibliographystyle{acm}
\bibliography{bibliography}}

\section*{Appendix A: NetFilter rules}

{\tt \small
\begin{verbatim}
iptables -t nat -A POSTROUTING -o eth0 -j MASQUERADE
iptables -A FORWARD -i eth0 -o eth1 -m state \
    --state NEW,RELATED,ESTABLISHED -j ACCEPT
iptables -A FORWARD -i eth0 -o eht1 -j ACCEPT
\end{verbatim}
}

\section*{Appendix B: Person-hours to verify \vignat}

Table~\ref{tab:man-hours} shows the number of person-hours spent to

\begin{table*}[h!]
\centering
\caption{Developer person-hours spent on each component part in hours.}
\label{tab:man-hours}
\begin{tabular}{lllllll}
\hline
               & Interface design & Implementation & Formal spec & Symbolic model & Verification & Validator plug-in \\ \hline
Hash Map        & 2.5*       & 2*             & 4.5*                 & 5.5*           & 73.5         & 11.5             \\
Port Allocator & 1*         & 1.5*           & 3.5*                 & 1.5*           & 70           & 7.5              \\
Static Array   & 1.5        & 0.5            & 2.5                  & 2              & 6            & 2                \\
Batcher        & 1          & 0.2            & 0.5                  & 1              & 0.5          & 0.2              \\ \hline
\end{tabular}
\end{table*}

\end{document}